---
title: "Tooth Growth EDA & Inference — Project Log"
author: "Brian Davies"
date: "19 September 2014"
output: pdf_document
---

## Data Access, dataset summary, EDA

```{r data_access_and_summary, cache = TRUE }
data( "ToothGrowth" )
head(ToothGrowth,n=3)
table(ToothGrowth$supp,ToothGrowth$dose)
```
60 data points, evenly distributed between two `supp`lements
{VC, OJ} (_Vitamin C_ and _Orange Juice_) and three `dose` levels,
{0.5, 1.0 and 2.0} -- no controls at `dose` = 0. So
6 combinations of `supp` and `dose` with 10 tests for each. No
natural pairing, so each set of ten
tests is a separate sample.

```{r EDA_pic, cache = TRUE, warning = FALSE, message = FALSE, fig.height = 2 }
library( "ggplot2" )
qplot( dose, len, data=ToothGrowth, colour = supp ) + geom_smooth()
```

Looks like a pretty clear increasing relationship, with `len` for `supp == 'OJ'` > `len`
for `supp == 'VC'` at the lower end, converging at the higher end, also some
saturation setting in for `dose > 1.0` for `supp == OJ`

## Hypotheses to be tested

We want to do multiple tests & apply a multi-comparison
correction, so it makes sense to marshall all our null
hypotheses in one place. They are the negations of the "gut feel"
for "how these things out to work", namely:

(1) Higher `dose` is associated with
higher levels of `len`. Four 1-sided T tests.
(2) At each `dose` level, changing the `supp` changes the
level of `len`. Three two-sided T tests.

## T-test comparison

### Compare adjacent dose levels for each supplement (four one-sided tests)

```{r compare_adjacent_dose_levels, cache = TRUE}
for( supp in c( 'OJ', 'VC' ) ){
  for ( dPair in list( c(0.5, 1.0), c(1.0, 2.0 ) ) ){
    thisT <- t.test( ToothGrowth[((ToothGrowth$supp==supp)&(ToothGrowth$dose==dPair[1])),'len'],
    ToothGrowth[((ToothGrowth$supp==supp)&(ToothGrowth$dose==dPair[2])),'len'],
    alternative = 'l' )
    message( sprintf( "## Results: supp = %2s, dose = %.1f & %.1f, p-value = %.5f, null-hypothesis CI = (%.2f, %.2f)",
    supp, dPair[1], dPair[2], thisT$p.value, thisT$conf.int[1], thisT$conf.int[2] ) )
  }
}
```

### Compare supplements at each dose level (three two-sided tests)

```{r compare_OJ_and_VC_at_each_dose, cache = TRUE}
for( dose in c(0.5,1.0,2.0)){
  thisT <- t.test( ToothGrowth[((ToothGrowth$supp=='OJ')&(ToothGrowth$dose==dose)),'len'],
    ToothGrowth[((ToothGrowth$supp=='VC')&(ToothGrowth$dose==dose)),'len'])
  message( sprintf( "## Results: dose = %.1f, p-value = %.5f, null-hypothesis CI = (%.2f, %.2f)",
    dose, thisT$p.value, thisT$conf.int[1], thisT$conf.int[2] ) ) }
```
