---
title: "Tooth Growth EDA & Inference — Project Log"
author: "Brian Davies"
date: "19 September 2014"
output: html_document
---

## Data Access and EDA

We're obviously supposed to figure out whether `len` depends on `dose` and
whether the dependency further depends on `supp`. Former is a real variable, latter
is a categorical variable, so plot `len` against `dose` for each value of `supp`

```{r data_access_and_EDA, cache = TRUE, warning = FALSE, message = FALSE }
data( "ToothGrowth" )
library( "ggplot2" )
qplot( dose, len, data=ToothGrowth, colour = supp, size = 2 ) + geom_smooth( size = 1 )
```

Looks like a pretty clear increasing relationship, with `len` for `supp` == OJ > `len`
for `supp` == VC at the lower end, converging at the higher end, also some
saturation setting in for `dose` > 1.0 for `supp` == OJ

## Summarizing the dataset

```{r ToothGrowth_dataset_summary, cache = TRUE }
table(ToothGrowth$supp,ToothGrowth$dose)
```

Data points are evenly distributed by supplement-type and dose
